RVD2: An ultra-sensitive variant detection model for low-depth heterogeneous next-generation sequencing data

### 2015-03-09 doc/bioinformatics2014/revision2/response_to_reviews.txt

Section 4.1 describing the details of the experimental protocol for the synthetic sequencing data is indeed brief. We have added the details the reviewer suggested and as expected, the narrative is much clearer. Briefly, Supplementary Table S1 in the Nucleic Acid Research publication shows that each sample has roughly 1,000,000 35bp paired-end reads with no errors in the multiplexing barcodes to give an average read depth of ~100,000 across the synthetic amplicon. We describe this and more detail as suggested in the protocol section.

5. The authors greatly expose the comparison of sensibility, specificity, true positive and true negative rates. It seems that the caveat of the method is on memory consumption and
processing time. It would be very informative to explicitly have a table with execution time
and memory consumption of all utilized methods, preferably in a more realistic scenario,
more specifically, a whole genome analysis.

The memory consumption and computational time for the algorithm is an important consideration. We intentionally focused on the accuracy of the methodology here and so chose Metropolis-Hastings sampling as our inference algorithm. Metropolis-Hastings is known to converge to sampling from the true posterior distribution in the limit. As such, presently the algorithm is not as computationally fast as it could be. To improve the speed of the algorithm, we have implemented a multi-threaded version to take advantage of a multi-core computational configuration. We provide computational speed results in Section X.X and discuss avenues for improvement that we are currently pursuing.

Reviewer 2:

3. On figure 6 panel A; at position index 35, 36, 37 and position 50; please check that the symbols shown for miscall versus mutant calls are as you intend them.. looks like two are reversed.

We are unable to identify what symbols are reversed. The RVD2 germline and somatic calls in Figure 6 do correspond to the same calls in Figure 5. Figure 5 and Figure 6 do use different symbols to denote their data, but the data is consistent.

### 2015-03-09 doc/bioinformatics2014/revision2/response_to_reviews.txt

5. The authors greatly expose the comparison of sensibility, specificity, true positive and true negative rates. It seems that the caveat of the method is on memory consumption and
processing time. It would be very informative to explicitly have a table with execution time
and memory consumption of all utilized methods, preferably in a more realistic scenario,
more specifically, a whole genome analysis.

The memory consumption and computational time for the algorithm is an important consideration. We intentionally focused on the accuracy of the methodology here and so chose Metropolis-Hastings sampling as our inference algorithm. Metropolis-Hastings is known to converge to sampling from the true posterior distribution in the limit. As such, presently the algorithm is not as computationally fast as it could be. To improve the speed of the algorithm, we have implemented a multi-threaded version to take advantage of a multi-core computational configuration. We provide computational speed results in Section X.X and discuss avenues for improvement that we are currently pursuing.

Reviewer 2:

2. on preprocessing methods (4.1); please check what you mean by the C50 option of BWA.. That should be something else.

Thank you for catching this important detail in the description of our methodology. We 
used the aligned bam files described previously, and ran samtools mmpileup with the -C50 flag to filter for high quality reads. The -C50 options is recommended in the samtools manual: http://samtools.sourceforge.net/mpileup.shtml. We have corrected the manuscript to reflect this.

### 2015-03-09 doc/bioinformatics2014/revision2/document.tex
Respond to reviewer 2 comment 2 regarding -C50 option and correct text in manuscript.

\subsubsection*{Pre-processing Methods.}
 We aligned the reads to the reference sequence using BWA v0.7.4 with the -C50 option to filter for high mapping quality reads. To simulate lower coverage data while retaining the error structure of real NGS data, BAM files for the synthetic DNA data were downsampled $10\times$, $100\times$, $1,000\times$, and $10,000\times$ using Picard v1.96. The final data set contains read pairs for three replicates of each case and pairs of reads three replicates for the control sample giving $N=6$ replicates for the control and each MAF level.

>>>>

\subsubsection*{Pre-processing Methods.}
The reads were aligned with Eland as described previously. We then ran samtools mpileup with the -C50 option to filter for high mapping quality reads. To simulate lower coverage data while retaining the error structure of real NGS data, BAM files for the synthetic DNA data were downsampled $10\times$, $100\times$, $1,000\times$, and $10,000\times$ using Picard v1.96. The final data set contains read pairs for three replicates of each case and pairs of reads three replicates for the control sample giving $N=6$ replicates for the control and each MAF level.
